Illuminating a Dark Lens : A Type Ia Supernova Magnified by \\ the Frontier Fields Galaxy Cluster Abell 2744

### 2015-08-10 APJ.submitted/snTomas.tex
Updated with response to referee. Ready for resubmission

\title{Illuminating a Dark Lens : A Type Ia Supernova Magnified by \\ the \changeR{Frontier Fields} Galaxy Cluster Abell 2744}

@@

In the case of lensed SNe, we can also reverse the experimental setup:
instead of using strong-lensing clusters to study distant SNe, we can
use the SNe as tools for examining the lenses \citep{Riehm:2011}.  The
most valuable transients for testing and improving cluster lens models
would be {\it strongly lensed} SNe that are resolved into multiple
images\ \citep{Holz:2001,Oguri:2003}.  Transients that are lensed into
multiple images can also become cosmological tools, as the measurement
of time delays between the images can provide cosmographic information
to constrain the Hubble parameter \citep{Refsdal:1964} and other
cosmological parameters \citep{Linder:2011}.  We have recently
observed the first example of a multiply-imaged SN \citep{Kelly:2015}.
We expect to detect the reappearance of this object (called ``SN
Refsdal'') \change{within the next
year} \citep{Oguri:2015,Sharon:2015,Diego:2015}, delivering a precise
test of lens model predictions.  Although detections of such objects
are currently very unlikely \citep{Li:2012}, they will become much
more common in the next decade \citep{Coe:2009,Dobke:2009}, and may be
developed into an important new cosmological
tool \citep{Oguri:2010,Linder:2011}.

In addition to time delays from multiply-imaged SNe, we can also
put cluster mass models to the test with the much more common category
of Type Ia SNe that are magnified but not multiply-imaged.  
\citet[][hereafter P14]{Patel:2014} and \citet{Nordin:2014} presented
independent analyses of three lensed SNe, of which at least 2 are
securely classified as Type Ia SNe -- all found in the Cluster Lensing
and Supernova survey with Hubble (CLASH, PI:Postman, HST Program ID
12068, \citealt{Postman:2012}).  Both groups demonstrated that these
standard candles can be used to provide accurate and precise
measurements of the true absolute magnification along a random sight
line through the cluster.  Although in these cases the SNe were used
to {\it test} the cluster mass models, one could in principle
incorporate the measured magnifications of Type Ia SNe into the
cluster as additional model constraints.  In that role, Type Ia SNe
have the particular value that they can be found anywhere in the
cluster field.  Thus, they can deliver model constraints in regions of
``middle distance'' from the cluster core, where both strong- and
weak-lensing constraints are unavailable. \change{Moreover, given
enough time, multiple background SNe Ia could be measured behind the
same cluster, each providing a new magnification constraint.}

One of the key values in observing standard candles behind
gravitational lenses is in addressing the problem of the mass-sheet
degeneracy \citep{Falco:1985,Schneider:1995}.  This degeneracy arises
because one can introduce into a lens model an unassociated sheet of
uniform mass in front of or behind the lens, without disturbing the
primary observable quantities.  For example, take a lens model with a
given surface mass density $\kappa$, and then transform the surface
mass density to $\kappa^{\prime}=(1-\lambda)\kappa+\lambda$ for any
arbitrary value $\lambda$.  Both the $\kappa$ and $\kappa^{\prime}$
models will produce exactly the same values for all positional and
shear constraints from strong and weak lensing \citep{Seitz:1997}.
When lensed background sources are available across a wide range of
redshifts (as is the case for the Abell 2744 cluster discussed here),
it should in principle be possible to break this
degeneracy \citep{Seitz:1997,Bradac:2004}.  However, there are more
complex versions of positional constraint
degeneracies \citep{Liesenborgs:2012,Schneider:2014}.  Such
degeneracies do not extend to the {\it absolute} magnification of a
background source's flux and size. \change{In the case of the simple
mass sheet degeneracy described above, the magnification scales as
$\mu\propto(1-\lambda)^{-2}$ \citep[see e.g.][]{Bartelmann:2010}.}
Therefore, one can break these fundamental degeneracies with an
absolute measurement of magnification from a standard
candle \citep{Holz:2001} or a standard ruler \citep{Sonnenfeld:2011}.

@@

\begin{figure*}
\begin{center}
%\includegraphics[width=\textwidth]{FIG/discovery_image}
\includegraphics[width=\textwidth]{FIG/discovery_image_lowres}
\caption{  \label{fig:DiscoveryImage} 
SN \tomas\ in the Abell 2744 field.  The left panel shows a
UV/Optical/IR color composite image constructed from all available HST
imaging of the Abell 2744 cluster field.  The inset panels on the right
show F814W imaging of the immediate vicinity of SN \tomas,
approximately 40\arcsec\ from the center of the cluster. The top panel
shows the template image, combining all data prior to the SN
appearance.  Labeled ellipses mark the nearest galaxies and their
spectroscopic redshift constraints, with the most likely host
galaxy marked in blue, and a background galaxy in yellow. The bottom
panel is constructed from all HFF F814W imaging taken while the SN was
detectable, and marks the SN location with an arrow.  (Left panel
image credit: NASA, ESA, and J. Lotz, M. Mountain, A. Koekemoer, and
the HFF Team (STScI)) }
\end{center}
\end{figure*}

@@

\input{TABLE/positions}

\input{TABLE/photometry}

@@

\change{
The galaxy identified as the host is not close enough to the cluster
core to {\it necessarily} be multiply-imaged, but it is still possible
that the host galaxy is one of the outer images of a multiple image
system.  In such a case, and given the position of the SN host, one
would expect that another image of the galaxy would be present at a
similar brightness, and would therefore be detectable in HST
imaging. To date, no plausible candidate for a counter-image has been
identified.}

\change{To measure the stellar mass of the host galaxy
we use Eq. 8 of \citet{Taylor:2011}, which relates the rest-frame
(g-i) color and i-band luminosity to the total stellar mass. To derive
these values, we fixed the redshift at $z=1.3457$ and repeated the SED
fitting using BPZ. From the best-fit SED we extracted rest-frame
optical magnitudes, and corrected them for lensing using a
magnification factor of $\mu=2.0$ -- a value that we will derive from
the SN itself in Section~\ref{sec:DistanceAndMagnification}.  From
this we determine the host galaxy mass to be 
\changeR{$10^{9.8}$ \Msun.} }

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/specfit}
\caption{  \label{fig:SpecFit}
Redshift and age determination from spectral template matching to the
the SN \tomas\ maximum light spectrum.  The y axis plots flux in
arbitrary units, and the x axis marks wavelength in \AA\ with the
observer-frame on the bottom and rest-frame on the top.  The \tomas\
spectrum observed with the HST ACS G800L grism is shown in black,
overlaid with model fits derived from a library of Type Ia templates
that have extended rest-frame UV coverage.  The top two panels show
the best matching templates when using a smooth 3rd-order polynomial
to warp the shape of the template pseudo-continuum, with the redshift
fixed at z=1.3457 and then allowed to float as a free parameter.  The
lower three panels show matches found when the templates are not
warped, both with and without fixing the redshift.  The bottom panel
shows the best match in this set, although at $z=0.98$ it is
inconsistent with the host galaxy redshift prior and the light curve.
}
\end{center}
\end{figure}

A spectrum of SN \tomas\ was collected with the ACS G800L grism on
2014 June 4 and 7, when the SN was \change{within 3 observer-frame
days of the observed peak brightness in the F814W band.}  The
observations -- listed at the bottom of Table~\ref{tab:Photometry} --
used 5 HST orbits from the FrontierSN program for a total
spectroscopic exposure time of $\sim$10 ksec.  The grism data were
processed and the target spectrum was extracted using a custom
pipeline \citep{Brammer:2012}, which was developed for the 3D-HST
program (PI:Van Dokkum; PID:12177, 12328) and also used by the GLASS
team.

@@

From the preceding analysis, we conclude that \tomas\ is a Type Ia SN
at $z\approx1.35$.  At this redshift, the absorption at $\sim$8700\AA\
corresponds to the blended \ion{Ca}{2} H\&K \change{and \ion{Si}{2} $\lambda$3858} features.
\change{Generally referred to as the Ca H\&K feature, this} absorption is commonly seen in Type Ia SN spectra
near maximum light, although it is also prominent in the spectra of
Type Ib and Ic core collapse SNe (\CCSNe).  The red color of
the \tomas\ SED is qualitatively consistent with a redshift of $z>1$
-- although this information was not used by {\tt SNID} for the
template matching.  As we will see in
Section~\ref{sec:PhotometricClassification}, this spectral
classification of SN \tomas\ is reinforced by the photometric
information, which also supports classification as a Type Ia SN at
$z\approx1.35$.

@@

To refine the redshift and phase constraints on \tomas, we next fit
the spectrum with a custom spectral matching program that employs a
library of Type Ia SN SEDs.  This library is similar to the Type Ia
spectral set used by SNID, but also includes more recent SNe with
well-observed spectral time series that extend to rest-frame UV
wavelengths (e.g. SN 2011fe and 2014J).  We first use an approach
similar to the SNID algorithm: warping the pseudo-continuum of each
template spectrum by dividing out a 3$^{\rm rd}$-order polynomial to
match the observed SED of SN \tomas.  This approach will find
templates that have similar abundances and photospheric
velocities. With the redshift fixed at $z=1.3457$ we find the best fit
is a spectrum from the normal Type Ia SN 2014J \citep{Foley:2014},
with a $\chi^2$ per degree of freedom $\nu$ of $\chi^2/\nu=121.8/89$
shown in the top panel of Figure~\ref{fig:SpecFit}. \change{The excess
variance in this fit may be attributed to the intrinsic variation of
Type Ia SN spectra, which is more prominent at UV
wavelengths \citep[e.g.][]{Foley:2008b,Wang:2012} and is not fully
represented in the available template library.}  Allowing the redshift
as an additional free parameter, we still find results that are
consistent with the SNID fits and $<2\sigma$ from the host galaxy
redshift: $z=1.31\pm0.02$ and a phase of $0\pm3$ rest-frame days. The
best-fitting spectral template in this case is the normal Type Ia SN
2012cg \citep{Amanullah:2015} at $z=1.31$ with $\chi^2/\nu=98.5/88$
(second panel of Figure~\ref{fig:SpecFit}).

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_lightcurve_classification}
\caption{ 
Maximum likelihood model for each SN sub-class, derived from Bayesian
model selection using the photometric data alone. Grey points show the
observed SN \tomas\ photometry with error bars, though these are
typically smaller than the size of the marker.  The Type Ia model
(orange solid line) is drawn from the SALT2 template at $z=1.35$. The
best match from all Type Ib/Ic models is based on the Type Ic SN
SDSS-14475 at $z=0.695$ (green dashed line). For the Type II class,
the best match is from the Type II-L SN 2007pg at $z=1.8$ (blue
dash-dot line). The Type Ia model is by far the best match, and the
only one that is consistent with both the photo-$z$ of the probable
host galaxy and the spectroscopic redshift from the SN spectrum.  The
date of the HST spectral observations is marked with a thin grey
vertical line.
\label{fig:photoclass} }
\end{center}
\end{figure}

As a check on the spectral classification of SN \tomas\
(Section~\ref{sec:SNID}), we independently classified the SN using a
Bayesian photometric classifier.  We use the {\tt sncosmo} software
package\footnote{\url{http://sncosmo.github.io/}} to simulate SN light
curves from $z=0.3$ to 2.3 and evaluate the classification probability
using traditional Bayesian model selection \citep[as
in][]{Jones:2013,Rodney:2014,Graur:2014,Rodney:2015a}.  In this
analysis we represent normal Type Ia SNe with the SALT2
model \citep{Guy:2010}, and \CCSNe\ with 42 discrete templates (26
Type II and 16 Type Ib/c) drawn from the template library of the
SuperNova Analysis software
package \citep[SNANA,][]{Kessler:2009a}.\footnote{Throughout this work
we use SNANA v10\_35g.}  Likelihoods are defined by comparing the
observed fluxes to model predictions in all passbands where the model
is defined.  In practice, this means we exclude the SN detections in
the F435W and F606W bands, which are too blue for our models at
$z>0.85$.

The \CCSN\ models have free parameters for date of peak brightness
($t_{\rm pk}$), amplitude, and redshift ($z$). Due to the expected
impact of gravitational lensing magnification, we do not include any
prior on the intrinsic luminosity for any SN sub-class.  We also do
not assign a prior for the SN redshift. This allows our photometric
analysis to provide an independent check on the host galaxy photo-$z$
and spectroscopic redshift (Sections~\ref{sec:DiscoveryAndFollowup}
and \ref{sec:SpecRedshift}).

@@

With the type and redshift securely defined as a normal Type Ia SN at
$z=1.3457$, we now turn to fitting the light curve with Type Ia
templates to measure the distance
modulus \change{(\S\ref{sec:LightCurveFitting})} and then derive the
gravitational lensing
magnification \change{(\S\ref{sec:ControlSampleComparison})}.  In this
process, we would like to avoid introducing systematic uncertainties
inherent to any assumed cosmological model.  To that end, \change{in
section \ref{sec:ControlSampleComparison}} we
follow \citealt{Patel:2014} and define the magnification by comparing
the measured distance modulus of SN \tomas\ against an average
distance modulus derived from a ``control sample'' of unlensed Type Ia
SNe at similar redshift. This allows us to make only the minimal
assumption that the redshift-distance relationship for Type Ia SN is
smooth and approximately linear over a small redshift span, which
should be true for any plausible cosmological model.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{FIG/snTomas_lightcurve_fit_fluxAB25}
\caption{ Type Ia light curve fits to SN \tomas\ using the MLCS2k2
(top row) and SALT2 (bottom row) fitters. The model redshifts are set to
$z=1.3457$ as determined from combined
spectroscopic and photometric constraints.  Solid lines denote the
best-fit model and shaded lines show the range allowed by 1-$\sigma$
uncertainties on the model parameters. 
Observed fluxes are shown as diamonds, scaled to an AB magnitude
zero point of 25. Error bars are plotted, but most are commensurate
with the size of the points. The left-most panel includes observations
in the F435W and F606W filters, although these were not used for the
fit, as they are bluer than the minimum wavelength for the model. The
lower axis marks time in observer frame days, while the top axis shows
the time in the rest frame relative to the epoch of peak brightness.
\label{fig:LightCurveFits} }
\end{center}
\end{figure*}

@@

As in \citealt{Patel:2014}, we derive a distance modulus for
SN \tomas\ and all SNe in our control sample using two independent
light curve fitters: the SALT2 model described above and the MLCS2k2
model \citep{Jha:2007}.    With both fitters we find light curve
shape and color parameters for \tomas\ that are fully consistent with
a normal Type Ia SN.  For SALT2, with the
redshift fixed at $z=1.3457$, we find a light curve shape
parameter of $x_1=0.135\pm0.199$ and a color parameter of
$c=-0.127\pm0.025$, yielding a $\chi^2$ value of 45.0 for 36 degrees
of freedom, $\nu$.  With the MLCS2k2 fitter the best-fit shape
parameter is $\Delta=-0.082\pm0.070$ and the color term is
$A_{\rm V}=0.011\pm0.025$, giving $\chi^2/\nu=22.8/36$.

@@

\input{TABLE/dist_mag}

The SNANA version of the MLCS2k2 fitter returns a value for the
distance modulus (dm$_{\rm MLCS2k2}$) that has an arbitrary zero point
offset relative to the SALT2 distances (dm$_{\rm SALT2}$).  To put the
two distances onto the same reference frame we add a zeropoint
correction of 0.20 mag to the MLCS2k2 distances as
in \citealt{Patel:2014}.  This correction was derived by applying both
fitters to a sample of Type Ia SNe from the SDSS
survey \citep{Holtzman:2008,Kessler:2009b}, with the extinction law
$R_V$ fixed at 1.9.

@@

It is now an accepted practice in cosmological analyses using Type Ia
SN to apply a correction to the luminosity of each SN based on the
stellar mass of its host galaxy. Typically this is described as a
simple bifurcation of the SN population: SNe that appear in more
massive hosts are observed to be $\sim$0.08 mag brighter (after
corrections for light curve shape and color) than SNe in low-mass
hosts \citep{Kelly:2010,Sullivan:2010}.  The dividing line for this
purely empirical ``mass step'' correction is generally set around
$10^{10}$\Msun.  \changeR{Although this threshold value} is somewhat arbitrary \citep[see
e.g.][]{Betoule:2014}, \changeR{it happens to be very close to the
-SN \tomas\ host galaxy mass of $10^{9.8}$\Msun (Section~\ref{sec:HostGalaxy}).}

@@

\changeR{
If we were to apply the correction to this sample using the standard
approach, then the SN \tomas\ distance modulus would not be adjusted,
because its host galaxy mass (Section~\ref{sec:HostGalaxy}) is just
below 10$^{10}$ \Msun. Note that there is some circularity here, as we
used the SN magnification in deriving the host galaxy mass, and now
use the host mass to inform the magnification.  However, regardless of
whether the \tomas\ host mass falls above the mass threshold or below,
this correction is not substantial enough to account for the observed
magnification tension.  The appropriate correction to be applied to
a SN above the mass threshold in this redshift range would be $<$0.04
mag \citep{Rigault:2013,Childress:2014b}, which results in a change to
the inferred magnification of \tomas\ that is much less than the
observed discrepancy of $\sim$0.24 mag.}

@@

\noindent The inferred magnifications for the two different fitters are
reported in the final column of Table~\ref{tab:MeasuredMagnification}.
\changeR{
Although we have chosen to use a cosmology-independent approach to
determine the magnifications, we note that these results are fully
consistent with the values that would be determined by comparing the
observed SN \tomas\ distance modulus against the predicted value from
a flat \LCDM\ cosmology. Adopting the cosmological parameters used
in \citetalias{Patel:2014} and derived in \citet{Sullivan:2011}
(\Ho=71.6, \Om=0.27, \OL=0.73) would give a predicted distance modulus
$dm_{\Lambda CDM}=44.90$ and a measured magnification $\mu=1.9$.
Substituting alternative cosmological
parameters \citep[e.g.,][]{Betoule:2014,Planck:2015} would not
significantly change the inferred magnification or affect our
conclusions.
}

>>>>

\title{Illuminating a Dark Lens : A Type Ia Supernova Magnified by \\ the Frontier Fields Galaxy Cluster Abell 2744}

@@

In the case of lensed SNe, we can also reverse the experimental setup:
instead of using strong-lensing clusters to study distant SNe, we can
use the SNe as tools for examining the lenses \citep{Riehm:2011}.  The
most valuable transients for testing and improving cluster lens models
would be {\it strongly lensed} SNe that are resolved into multiple
images\ \citep{Holz:2001,Oguri:2003}.  Transients that are lensed into
multiple images can also become cosmological tools, as the measurement
of time delays between the images can provide cosmographic information
to constrain the Hubble parameter \citep{Refsdal:1964} and other
cosmological parameters \citep{Linder:2011}.  We have recently
observed the first example of a multiply-imaged SN \citep{Kelly:2015}.
We expect to detect the reappearance of this object (called ``SN
Refsdal'') within the next
year \citep{Oguri:2015,Sharon:2015,Diego:2015}, delivering a precise
test of lens model predictions.  Although detections of such objects
are currently very unlikely \citep{Li:2012}, they will become much
more common in the next decade \citep{Coe:2009,Dobke:2009}, and may be
developed into an important new cosmological
tool \citep{Oguri:2010,Linder:2011}.

In addition to time delays from multiply-imaged SNe, we can also
put cluster mass models to the test with the much more common category
of Type Ia SNe that are magnified but not multiply-imaged.  
\citet[][hereafter P14]{Patel:2014} and \citet{Nordin:2014} presented
independent analyses of three lensed SNe, of which at least 2 are
securely classified as Type Ia SNe -- all found in the Cluster Lensing
and Supernova survey with Hubble (CLASH, PI:Postman, HST Program ID
12068, \citealt{Postman:2012}).  Both groups demonstrated that these
standard candles can be used to provide accurate and precise
measurements of the true absolute magnification along a random sight
line through the cluster.  Although in these cases the SNe were used
to {\it test} the cluster mass models, one could in principle
incorporate the measured magnifications of Type Ia SNe into the
cluster as additional model constraints.  In that role, Type Ia SNe
have the particular value that they can be found anywhere in the
cluster field.  Thus, they can deliver model constraints in regions of
``middle distance'' from the cluster core, where both strong- and
weak-lensing constraints are unavailable. Moreover, given
enough time, multiple background SNe Ia could be measured behind the
same cluster, each providing a new magnification constraint.

One of the key values in observing standard candles behind
gravitational lenses is in addressing the problem of the mass-sheet
degeneracy \citep{Falco:1985,Schneider:1995}.  This degeneracy arises
because one can introduce into a lens model an unassociated sheet of
uniform mass in front of or behind the lens, without disturbing the
primary observable quantities.  For example, take a lens model with a
given surface mass density $\kappa$, and then transform the surface
mass density to $\kappa^{\prime}=(1-\lambda)\kappa+\lambda$ for any
arbitrary value $\lambda$.  Both the $\kappa$ and $\kappa^{\prime}$
models will produce exactly the same values for all positional and
shear constraints from strong and weak lensing \citep{Seitz:1997}.
When lensed background sources are available across a wide range of
redshifts (as is the case for the Abell 2744 cluster discussed here),
it should in principle be possible to break this
degeneracy \citep{Seitz:1997,Bradac:2004}.  However, there are more
complex versions of positional constraint
degeneracies \citep{Liesenborgs:2012,Schneider:2014}.  Such
degeneracies do not extend to the {\it absolute} magnification of a
background source's flux and size. In the case of the simple
mass sheet degeneracy described above, the magnification scales as
$\mu\propto(1-\lambda)^{-2}$ \citep[see e.g.][]{Bartelmann:2010}.
Therefore, one can break these fundamental degeneracies with an
absolute measurement of magnification from a standard
candle \citep{Holz:2001} or a standard ruler \citep{Sonnenfeld:2011}.

@@

\begin{figure*}
\begin{center}
%\includegraphics[width=\textwidth]{discovery_image}
\includegraphics[width=\textwidth]{discovery_image_lowres}
\caption{  \label{fig:DiscoveryImage} 
SN \tomas\ in the Abell 2744 field.  The left panel shows a
UV/Optical/IR color composite image constructed from all available HST
imaging of the Abell 2744 cluster field.  The inset panels on the right
show F814W imaging of the immediate vicinity of SN \tomas,
approximately 40\arcsec\ from the center of the cluster. The top panel
shows the template image, combining all data prior to the SN
appearance.  Labeled ellipses mark the nearest galaxies and their
spectroscopic redshift constraints, with the most likely host
galaxy marked in blue, and a background galaxy in yellow. The bottom
panel is constructed from all HFF F814W imaging taken while the SN was
detectable, and marks the SN location with an arrow.  (Left panel
image credit: NASA, ESA, and J. Lotz, M. Mountain, A. Koekemoer, and
the HFF Team (STScI)) }
\end{center}
\end{figure*}

@@

\input{positions}

\input{photometry}

@@

The galaxy identified as the host is not close enough to the cluster
core to {\it necessarily} be multiply-imaged, but it is still possible
that the host galaxy is one of the outer images of a multiple image
system.  In such a case, and given the position of the SN host, one
would expect that another image of the galaxy would be present at a
similar brightness, and would therefore be detectable in HST
imaging. To date, no plausible candidate for a counter-image has been
identified.

To measure the stellar mass of the host galaxy
we use Eq. 8 of \citet{Taylor:2011}, which relates the rest-frame
(g-i) color and i-band luminosity to the total stellar mass. To derive
these values, we fixed the redshift at $z=1.3457$ and repeated the SED
fitting using BPZ. From the best-fit SED we extracted rest-frame
optical magnitudes, and corrected them for lensing using a
magnification factor of $\mu=2.0$ -- a value that we will derive from
the SN itself in Section~\ref{sec:DistanceAndMagnification}.  From
this we determine the host galaxy mass to be $10^{9.8}$ \Msun.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{specfit}
\caption{  \label{fig:SpecFit}
Redshift and age determination from spectral template matching to the
the SN \tomas\ maximum light spectrum.  The y axis plots flux in
arbitrary units, and the x axis marks wavelength in \AA\ with the
observer-frame on the bottom and rest-frame on the top.  The \tomas\
spectrum observed with the HST ACS G800L grism is shown in black,
overlaid with model fits derived from a library of Type Ia templates
that have extended rest-frame UV coverage.  The top two panels show
the best matching templates when using a smooth 3rd-order polynomial
to warp the shape of the template pseudo-continuum, with the redshift
fixed at z=1.3457 and then allowed to float as a free parameter.  The
lower three panels show matches found when the templates are not
warped, both with and without fixing the redshift.  The bottom panel
shows the best match in this set, although at $z=0.98$ it is
inconsistent with the host galaxy redshift prior and the light curve.
}
\end{center}
\end{figure}

A spectrum of SN \tomas\ was collected with the ACS G800L grism on
2014 June 4 and 7, when the SN was within 3 observer-frame
days of the observed peak brightness in the F814W band.  The
observations -- listed at the bottom of Table~\ref{tab:Photometry} --
used 5 HST orbits from the FrontierSN program for a total
spectroscopic exposure time of $\sim$10 ksec.  The grism data were
processed and the target spectrum was extracted using a custom
pipeline \citep{Brammer:2012}, which was developed for the 3D-HST
program (PI:Van Dokkum; PID:12177, 12328) and also used by the GLASS
team.

@@

From the preceding analysis, we conclude that \tomas\ is a Type Ia SN
at $z\approx1.35$.  At this redshift, the absorption at $\sim$8700\AA\
corresponds to the blended \ion{Ca}{2} H\&K and \ion{Si}{2} $\lambda$3858 features.  Generally referred to as the Ca H\&K feature, this absorption is commonly seen in Type Ia SN spectra
near maximum light, although it is also prominent in the spectra of
Type Ib and Ic core collapse SNe (\CCSNe).  The red color of
the \tomas\ SED is qualitatively consistent with a redshift of $z>1$
-- although this information was not used by {\tt SNID} for the
template matching.  As we will see in
Section~\ref{sec:PhotometricClassification}, this spectral
classification of SN \tomas\ is reinforced by the photometric
information, which also supports classification as a Type Ia SN at
$z\approx1.35$.

@@

To refine the redshift and phase constraints on \tomas, we next fit
the spectrum with a custom spectral matching program that employs a
library of Type Ia SN SEDs.  This library is similar to the Type Ia
spectral set used by SNID, but also includes more recent SNe with
well-observed spectral time series that extend to rest-frame UV
wavelengths (e.g. SN 2011fe and 2014J).  We first use an approach
similar to the SNID algorithm: warping the pseudo-continuum of each
template spectrum by dividing out a 3$^{\rm rd}$-order polynomial to
match the observed SED of SN \tomas.  This approach will find
templates that have similar abundances and photospheric
velocities. With the redshift fixed at $z=1.3457$ we find the best fit
is a spectrum from the normal Type Ia SN 2014J \citep{Foley:2014},
with a $\chi^2$ per degree of freedom $\nu$ of $\chi^2/\nu=121.8/89$
shown in the top panel of Figure~\ref{fig:SpecFit}. The excess
variance in this fit may be attributed to the intrinsic variation of
Type Ia SN spectra, which is more prominent at UV
wavelengths \citep[e.g.][]{Foley:2008b,Wang:2012} and is not fully
represented in the available template library.  Allowing the redshift
as an additional free parameter, we still find results that are
consistent with the SNID fits and $<2\sigma$ from the host galaxy
redshift: $z=1.31\pm0.02$ and a phase of $0\pm3$ rest-frame days. The
best-fitting spectral template in this case is the normal Type Ia SN
2012cg \citep{Amanullah:2015} at $z=1.31$ with $\chi^2/\nu=98.5/88$
(second panel of Figure~\ref{fig:SpecFit}).

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{snTomas_lightcurve_classification}
\caption{ 
Maximum likelihood model for each SN sub-class, derived from Bayesian
model selection using the photometric data alone. Grey points show the
observed SN \tomas\ photometry with error bars, though these are
typically smaller than the size of the marker.  The Type Ia model
(orange solid line) is drawn from the SALT2 template at $z=1.35$. The
best match from all Type Ib/Ic models is based on the Type Ic SN
SDSS-14475 at $z=0.695$ (green dashed line). For the Type II class,
the best match is from the Type II-L SN 2007pg at $z=1.8$ (blue
dash-dot line). The Type Ia model is by far the best match, and the
only one that is consistent with both the \change{spectroscopic
redshift} of the probable host galaxy and the spectroscopic redshift
from the SN spectrum.  The date of the HST spectral observations is
marked with a thin grey vertical line.
\label{fig:photoclass} }
\end{center}
\end{figure}

As a check on the spectral classification of SN \tomas\
(Section~\ref{sec:SNID}), we independently classified the SN using a
Bayesian photometric classifier.  We use the {\tt sncosmo} software
package\footnote{\url{http://sncosmo.github.io/}} to simulate SN light
curves from $z=0.3$ to 2.3 and evaluate the classification probability
using traditional Bayesian model selection \citep[as
in][]{Jones:2013,Rodney:2014,Graur:2014,Rodney:2015b}.  In this
analysis we represent normal Type Ia SNe with the SALT2
model \citep{Guy:2010}, and \CCSNe\ with 42 discrete templates (26
Type II and 16 Type Ib/c) drawn from the template library of the
SuperNova Analysis software
package \citep[SNANA,][]{Kessler:2009a}.\footnote{Throughout this work
we use SNANA v10\_35g.}  Likelihoods are defined by comparing the
observed fluxes to model predictions in all passbands where the model
is defined.  In practice, this means we exclude the SN detections in
the F435W and F606W bands, which are too blue for our models at
$z>0.85$.

The \CCSN\ models have free parameters for date of peak brightness
($t_{\rm pk}$), amplitude, and redshift ($z$). Due to the expected
impact of gravitational lensing magnification, we \change{use a flat
prior for the intrinsic luminosity of all SN sub-classes.  We also
apply a flat prior for the SN redshift, which} allows our photometric
analysis to provide an independent check on the host galaxy photo-$z$
and spectroscopic redshift (Sections~\change{\ref{sec:HostGalaxy}}
and \ref{sec:SpecRedshift}).

@@

With the type and redshift securely defined as a normal Type Ia SN at
$z=1.3457$, we now turn to fitting the light curve with Type Ia
templates to measure the distance
modulus (\S\ref{sec:LightCurveFitting}) and then derive the
gravitational lensing
magnification (\S\ref{sec:ControlSampleComparison}).  In this
process, we would like to avoid introducing systematic uncertainties
inherent to any assumed cosmological model.  To that end, in
section \ref{sec:ControlSampleComparison} we
follow \P14\ and define the magnification by comparing
the measured distance modulus of SN \tomas\ against an average
distance modulus derived from a ``control sample'' of unlensed Type Ia
SNe at similar redshift. This allows us to make only the minimal
assumption that the redshift-distance relationship for Type Ia SN is
smooth and approximately linear over a small redshift span, which
should be true for any plausible cosmological model.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{snTomas_lightcurve_fit_fluxAB25}
\caption{ Type Ia light curve fits to SN \tomas\ using the MLCS2k2
(top row) and SALT2 (bottom row) fitters. The model redshifts are set to
$z=1.3457$ as determined from combined
spectroscopic and photometric constraints.  Solid lines denote the
best-fit model and shaded lines show the range allowed by 1-$\sigma$
uncertainties on the model parameters. 
Observed fluxes are shown as diamonds, scaled to an AB magnitude
zero point of 25. Error bars are plotted, but most are commensurate
with the size of the points. The left-most panel includes observations
in the F435W and F606W filters, although these were not used for the
fit, as they are bluer than the minimum wavelength for the model. The
lower axis marks time in observer frame days, while the top axis shows
the time in the rest frame relative to the epoch of peak brightness.
\label{fig:LightCurveFits} }
\end{center}
\end{figure*}

@@

As in \P14, we derive a distance modulus for
SN \tomas\ and all SNe in our control sample using two independent
light curve fitters: the SALT2 model described above and the MLCS2k2
model \citep{Jha:2007}.    With both fitters we find light curve
shape and color parameters for \tomas\ that are fully consistent with
a normal Type Ia SN.  For SALT2, with the
redshift fixed at $z=1.3457$, we find a light curve shape
parameter of $x_1=0.135\pm0.199$ and a color parameter of
$c=-0.127\pm0.025$, yielding a $\chi^2$ value of 45.0 for 36 degrees
of freedom, $\nu$.  With the MLCS2k2 fitter the best-fit shape
parameter is $\Delta=-0.082\pm0.070$ and the color term is
$A_{\rm V}=0.011\pm0.025$, giving $\chi^2/\nu=22.8/36$.

@@

\input{dist_mag}

The SNANA version of the MLCS2k2 fitter returns a value for the
distance modulus (dm$_{\rm MLCS2k2}$) that has an arbitrary zero point
offset relative to the SALT2 distances (dm$_{\rm SALT2}$).  To put the
two distances onto the same reference frame we add a zeropoint
correction of 0.20 mag to the MLCS2k2 distances as
in \P14.  This correction was derived by applying both
fitters to a sample of Type Ia SNe from the SDSS
survey \citep{Holtzman:2008,Kessler:2009b}, with the extinction law
$R_V$ fixed at 1.9.

@@

It is now an accepted practice in cosmological analyses using Type Ia
SN to apply a correction to the luminosity of each SN based on the
stellar mass of its host galaxy. Typically this is described as a
simple bifurcation of the SN population: SNe that appear in more
massive hosts are observed to be $\sim$0.08 mag brighter (after
corrections for light curve shape and color) than SNe in low-mass
hosts \citep{Kelly:2010,Sullivan:2010}.  The dividing line for this
purely empirical ``mass step'' correction is generally set around
$10^{10}$\Msun.  Although this threshold value is somewhat arbitrary \citep[see
e.g.][]{Betoule:2014}, it happens to be very close to the
-SN \tomas\ host galaxy mass of $10^{9.8}$\Msun (Section~\ref{sec:HostGalaxy}).

@@

If we were to apply the correction to this sample using the standard
approach, then the SN \tomas\ distance modulus would not be adjusted,
because its host galaxy mass (Section~\ref{sec:HostGalaxy}) is just
below 10$^{10}$ \Msun. Note that there is some circularity here, as we
used the SN magnification in deriving the host galaxy mass, and now
use the host mass to inform the magnification.  However, regardless of
whether the \tomas\ host mass falls above the mass threshold or below,
this correction is not substantial enough to account for the observed
magnification tension.  The appropriate correction to be applied to
a SN above the mass threshold in this redshift range would be $<$0.04
mag \citep{Rigault:2013,Childress:2014b}, which results in a change to
the inferred magnification of \tomas\ that is much less than the
observed discrepancy of $\sim$0.24 mag.

@@

\noindent The inferred magnifications for the two different fitters are
reported in the final column of Table~\ref{tab:MeasuredMagnification}.
Although we have chosen to use a cosmology-independent approach to
determine the magnifications, we note that these results are fully
consistent with the values that would be determined by comparing the
observed SN \tomas\ distance modulus against the predicted value from
a flat \LCDM\ cosmology. Adopting the cosmological parameters used
in \P14\ and derived in \citet{Sullivan:2011}
(\Ho=71.6, \Om=0.27, \OL=0.73) would give a predicted distance modulus
$dm_{\Lambda CDM}=44.90$ and a measured magnification $\mu=1.9$.
Substituting alternative cosmological
parameters \citep[e.g.,][]{Betoule:2014,Planck:2015} would not
significantly change the inferred magnification or affect our
conclusions.
